Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-25 NEUROC/sec3-experiment-design.tex
sec3修改数据集的问题, sec4修改了inference分析的问题

We used six real-world graph datasets as listed in \tablename~\ref{tab:dataset_overview} that were popular in the GNN accuracy evaluation \cite{yang2016_revisiting_semisupervised, zeng2020_graphsaint, shchur2018_pitfall_of_gnn}.
%
For directed graphs, PyG converts them into undirected ones during data loading.
%
Thus, the average degree of a directed graph $\bar{d}=\frac{2|\mathcal{E}|}{|\mathcal{V}|}$.
%
For an undirected graph, the average degree is defined as $\bar{d}=\frac{|\mathcal{E}|}{|\mathcal{V}|}$.
%
For the \texttt{cam} dataset, we generated random dense feature vectors for it.
%
Since its vertices were not associated with feature vectors, we excluded it from accuracy evaluation.
%
We also used random graphs generated by the R-MAT graph generator \cite{rmat-generator} in some experiments, to explore the effects of graph topological characteristics (like average degrees) on performance bottlenecks.
%
Input feature vectors of random graphs were random dense vectors with the dimension of 32.
%
Vertices of random graphs were classified into 10 classes randomly.
%
For \texttt{amp}, \texttt{amc}, \texttt{cph}, and \texttt{cam}, we used 70\%/15\%/15\% of the samples as the training/evaluation/test set, respectively.
%
For \texttt{pub}, we used 500/1000 vertices as the evaluation/test set and the remained as the training set according to \cite{chen2018_fastgcn}.
%
For \texttt{fli}, we used 50\%/25\%/25\% of the samples as the training/evaluation/test set according to \cite{zeng2020_graphsaint}.

\begin{table}[H]
    \centering
    \begin{tabular}{cccccccc}
        \toprule
        Dataset                                                 & $|\mathcal{V}|$ & $|\mathcal{E}|$ & $\bar{d}$ & $dim(\boldsymbol{v})$ & \#Class & Directed \\
        \midrule
        pubmed (\texttt{pub}) \cite{yang2016_revisiting_semisupervised}  & 19,717          & 44,324          & 4.5       & 500                   & 3       & Yes      \\
        amazon-photo (\texttt{amp}) \cite{shchur2018_pitfall_of_gnn}     & 7,650           & 119,081         & 31.1      & 745                   & 8       & Yes      \\
        amazon-computers (\texttt{amc}) \cite{shchur2018_pitfall_of_gnn} & 13,752          & 245,861         & 35.8      & 767                   & 10      & Yes      \\
        coauthor-physics (\texttt{cph}) \cite{shchur2018_pitfall_of_gnn} & 34,493          & 247,962         & 14.4      & 8415                  & 5       & Yes      \\
        flickr (\texttt{fli}) \cite{zeng2020_graphsaint}                 & 89,250          & 899,756         & 10.1      & 500                   & 7       & No       \\
        com-amazon (\texttt{cam}) \cite{yang2012_defining}               & 334,863         & 925,872         & 2.8       & 32                    & 10      & No       \\
        \bottomrule
    \end{tabular}
    \caption{Dataset overview. $\bar{d}$ represents the average vertex degree. $dim(\boldsymbol{v})$ is the dimension of the input feature vector.}
    \label{tab:dataset_overview}
\end{table}

>>>>

We used six real-world graph datasets as listed in \tablename~\ref{tab:dataset_overview} that were popular in the GNN accuracy evaluation \cite{chen2018_fastgcn, zeng2020_graphsaint, shchur2018_pitfall_of_gnn}.
%
For directed graphs, PyG converts them into undirected ones during data loading.
%
Thus, the average degree of a directed graph $\bar{d}=\frac{2|\mathcal{E}|}{|\mathcal{V}|}$.
%
For an undirected graph, the average degree is defined as $\bar{d}=\frac{|\mathcal{E}|}{|\mathcal{V}|}$.
%
For the \texttt{cam} dataset, we generated random dense feature vectors for it.
%
Since its vertices were not associated with feature vectors, we excluded it from accuracy evaluation.
%
We also used random graphs generated by the R-MAT graph generator \cite{rmat-generator} in some experiments, to explore the effects of graph topological characteristics (like average degrees) on performance bottlenecks.
%
Input feature vectors of random graphs were random dense vectors with the dimension of 32.
%
Vertices of random graphs were classified into 10 classes randomly.
%
For \texttt{amp}, \texttt{amc}, \texttt{cph}, and \texttt{cam}, we used 70\%/15\%/15\% of the samples as the training/evaluation/test set, respectively.
%
For \texttt{pub}, we used 500/1000 vertices as the evaluation/test set and the remained as the training set according to \cite{chen2018_fastgcn}.
%
For \texttt{fli}, we used 50\%/25\%/25\% of the samples as the training/evaluation/test set according to \cite{zeng2020_graphsaint}.

\begin{table}[H]
    \centering
    \begin{tabular}{cccccccc}
        \toprule
        Dataset                                                 & $|\mathcal{V}|$ & $|\mathcal{E}|$ & $\bar{d}$ & $dim(\boldsymbol{v})$ & \#Class & Directed \\
        \midrule
        pubmed (\texttt{pub}) \cite{chen2018_fastgcn}  & 19,717          & 44,324          & 4.5       & 500                   & 3       & Yes      \\
        amazon-photo (\texttt{amp}) \cite{shchur2018_pitfall_of_gnn}     & 7,650           & 119,081         & 31.1      & 745                   & 8       & Yes      \\
        amazon-computers (\texttt{amc}) \cite{shchur2018_pitfall_of_gnn} & 13,752          & 245,861         & 35.8      & 767                   & 10      & Yes      \\
        coauthor-physics (\texttt{cph}) \cite{shchur2018_pitfall_of_gnn} & 34,493          & 247,962         & 14.4      & 8415                  & 5       & Yes      \\
        flickr (\texttt{fli}) \cite{zeng2020_graphsaint}                 & 89,250          & 899,756         & 10.1      & 500                   & 7       & No       \\
        com-amazon (\texttt{cam}) \cite{yang2012_defining}               & 334,863         & 925,872         & 2.8       & 32                    & 10      & No       \\
        \bottomrule
    \end{tabular}
    \caption{Dataset overview. $\bar{d}$ represents the average vertex degree. $dim(\boldsymbol{v})$ is the dimension of the input feature vector.}
    \label{tab:dataset_overview}
\end{table}

### 2020-12-25 NEUROC/sec4-experiments.tex
更新Reviewer2
sec3修改数据集的问题, sec4修改了inference分析的问题

We further investigated the most time-consuming step of the edge calculation stage.
%
In the implementation of PyG, the edge calculation stage consists of four steps: collection, messaging, aggregation, and vector updating, as shown in \figurename~\ref{fig:steps_in_edge_calculation}.
%
The edge index is a matrix with $|\mathcal{E}|$ rows and two columns.
%
It holds the edge set of the graph.
%
The two columns of the edge index store the source and the target vertex IDs of each edge.
%
The collection step copies the hidden vectors from the previous GNN layer $\MyVec{h}^l_x$ and $\MyVec{h}^l_y$ to the both endpoints of each edge $e_{x,y}$ in the edge index, forming the parameter tensor $[\MyVec{h}^l_x, \MyVec{h}^l_{y}, \MyVec{e}_{x,y}]$ of the messaging function $\phi^l$.
%
This step only involves data movement.
%
The messaging step calls the messaging function $\phi^l$ on all edges to get message vectors $\MyVec{m}^l_{x,y}$.
%
The aggregation step aggregates the message vectors with the same target vertex into an aggregated vector $\MyVec{s}^l_x$.
%
The vector updating step is optional.
%
It performs an additional transformation on the aggregated vectors (for example, adding the bias in GCN).

@@

% added by wangyunpan, 2020-12-22 ----start----
\paragraph{Inference Time Breakdown Analysis}
We conducted the same experiments in the inference stage as above.
%
The results of layer level experiment, step Level in edge Calculation experiment and operator Level experiment were in general consistent with those in the Training stage.
%
The performance bottleneck analysis in the Training stage can also be applied to the inference stage.
% added by wangyunpan, 2020-12-22 ----end----

@@

% 不同点: phase
\begin{figure}[H]
    \centering
    \includegraphics[height=5cm]{figs/experiments/exp_inference_full_memory_usage_stage_amp.pdf}
    \caption{Memory usage of each phase during the GNN training. Dataset: \texttt{amp}.}
    \label{fig:exp_inference_full_memory_usage_stage_amp}
\end{figure}

@@

% 不同点: input feature dims
\begin{figure}[H]
    \centering
    \includegraphics[height=5cm]{figs/experiments/exp_inference_full_memory_expansion_ratio_input_feature_dimension_com-amazon_expansion_ratio.pdf}
    \caption{Memory expansion ratio under different dimensions of input feature vectors. Dataset: \texttt{cam}.}
    \label{fig:exp_inference_full_memory_expension_ratio_input_feature_dimension}
\end{figure}

>>>>

We further investigated the most time-consuming step of the edge calculation stage.
%
In the implementation of PyG, the edge calculation stage consists of four steps: collection, messaging, aggregation, and vector updating, as shown in \figurename~\ref{fig:steps_in_edge_calculation}.
%
The edge index is a matrix with $|\mathcal{E}|$ rows and two columns.
%
It holds the edge set of the graph.
%
The two columns of the edge index store the source and the target vertex IDs of each edge.
%
The collection step copies the hidden vectors from the previous GNN layer $\MyVec{h}^l_y$ and $\MyVec{h}^l_x$ to the both endpoints of each edge $e_{y,x}$ in the edge index, forming the parameter tensor $[\MyVec{h}^l_{y}, \MyVec{h}^l_{x}, \MyVec{e}_{y,x}]$ of the messaging function $\phi^l$.
%
This step only involves data movement.
%
The messaging step calls the messaging function $\phi^l$ on all edges to get message vectors $\MyVec{m}^l_{y,x}$.
%
The aggregation step aggregates the message vectors with the same target vertex into an aggregated vector $\MyVec{s}^l_x$.
%
The vector updating step is optional.
%
It performs an additional transformation on the aggregated vectors (for example, adding the bias in GCN).

@@

% added by wangyunpan, 2020-12-22 ----start----
\paragraph{Inference Time Breakdown Analysis}
We conducted the same experiments in the inference stage as above.
%
The results of layer level experiment, step Level in edge calculation experiment and operator Level experiment were in general consistent with those in the Training stage.
%
The performance bottleneck analysis in the Training stage can also be applied to the inference stage.
% added by wangyunpan, 2020-12-22 ----end----

@@

% 不同点: phase
\begin{figure}[H]
    \centering
    \includegraphics[height=5cm]{figs/experiments/exp_inference_full_memory_usage_stage_amp.pdf}
    \caption{Memory usage of each phase during the GNN inference. Dataset: \texttt{amp}.}
    \label{fig:exp_inference_full_memory_usage_stage_amp}
\end{figure}

@@

% 不同点: input feature dims
\begin{figure}[H]
    \centering
    \includegraphics[height=5cm]{figs/experiments/exp_inference_full_memory_expansion_ratio_input_feature_dimension_com-amazon_expansion_ratio.pdf}
    \caption{Memory expansion ratio under different dimensions of input feature vectors during inference. Dataset: \texttt{cam}.}
    \label{fig:exp_inference_full_memory_expension_ratio_input_feature_dimension}
\end{figure}
